---
title: "Danish_Monarchs"
author: "Aske Høj Andersen, Vladimir Bountovski and Oscar Robert Nixon"
date: "2026-03-12"
output: html_document
---

```{r setup, include=TRUE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```


### Load in the Monarchs

```{r eval=TRUE}

#Testing different functions
monarchs1 <- read.csv("data/danish_monarchs.csv")

monarchs2 <- read_csv("data/danish_monarchs.csv")

monarchs3 <- read.csv2("data/danish_monarchs.csv")

monarchs4 <- read_csv2("data/danish_monarchs.csv")
```

1. Which of these functions is a `tidyverse` function? 
read_csv(): Tidyverse function used for reading standard comma-separated files.
read_csv2(): Tidyverse function used for reading semicolon-separated files.

2. What is the result of running `class()` on the `kings` object created with a tidyverse function.
[1] "spec_tbl_df" "tbl_df"      "tbl"         "data.frame" - identifies it as a tibble (tidyverse data frame).

3. How many columns does the object have when created with these different functions?
dim() gives me 53 collums

4. Show the dataset so that we can see how R interprets each column

```{r eval = TRUE}
monarchs <- read_csv("data/danish_monarchs.csv")

class(monarchs)

dim(monarchs)

glimpse(monarchs)

view(monarchs)
```


### Calculate the duration of reign for all danish kings/queens

```{r}
# YOUR CODE

monarchs_duration <- monarchs %>%
  mutate(durationreign=ReignEnd-ReignStart)

view(monarchs_duration)
glimpse(monarchs_duration)
tail(monarchs_duration)
```

### Calculating average reign for all rulers

```{r}
# YOUR CODE
duration <- as.numeric(monarchs_duration$durationreign)
class(duration)

mean(duration, na.rm = TRUE)
```

### Monarchs with longer than average reign

```{r}
# YOUR CODE
avg_duration <- mean(monarchs_duration$durationreign, na.rm = TRUE)

monarchs_filtered <- monarchs %>%
  filter(duration >= avg_duration)

monarchs_filtered

monarchs_filtered %>%
  count()

```


### Days of the longest three reigns

```{r}

longestreign <- monarchs %>%
  arrange(desc(monarchs_duration$durationreign)) %>% 
  slice(1:3)

monarchs_duration <- monarchs %>% 
  mutate(days=monarchs_duration$durationreign*365)

glimpse(monarchs_duration)
  
```


### Monarchs duration of reign through time

```{r}
# YOUR CODE

monarchs_duration %>%
  mutate(midyear = ReignStart + (ReignEnd - ReignStart)/2) %>%
  ggplot(aes(x = midyear, y = duration)) +
  geom_point() +
  geom_smooth()

```

